Models

### 2021-12-16 analysis/04_collate_rarefied_null.Rmd

Merge rarefied species and fd metrics and null models
```{r}
rarefied_metrics <- full_join(species_metrics, fd_metrics %>% select(-CWM),
                              c("YEAR", "cell", "rarefy_resamp", "rarefyID")) %>%
  full_join(cwm_metrics, c("YEAR", "cell", "rarefy_resamp", "rarefyID")) %>%
  filter(S != 1) %>% # remove samples that had one species as they result in strange metric values
  select(-qual.FRic)

@@

##	calculate the medians for all the metrics for studies that were rarefied
## exclude categorical variables - they don't make sense as medians
metric_cols <- ungroup(rarefied_metrics) %>%
  select(-c(SamplePool, SampleN, num_years, duration, startYear, endYear, type,
            diet_5cat, pelagicspecialist, diurnal, nocturnal, crepuscular, dir, lar)) %>%
  pivot_longer(-c(rarefyID, rarefied, YEAR, cell, rarefy_resamp), names_to = "metric") %>%
  left_join(null_table %>% select(-c(richness, se, lowerCI, upperCI)),
            by = c("YEAR" = "year", "rarefyID" = "rarefyid", "rarefy_resamp" = "rarefy_resamp", "metric" = "metric"))

@@

# calculate SES and incorporate into metric columns
metric_cols <- metric_cols %>%
  mutate(SES = (value - mean)/sd) %>%
  #for years where species richness = richness of the species pool, FRic sd = 0, so use the original value with no SES correction
  mutate(SES = case_when(
    sd == 0 ~ 0,
    TRUE ~ SES)) %>%
  select(-c(mean, sd)) %>%
  #the next three lines go back to wide format for both SES and unadjusted values, then to long format incorporating SES values
  pivot_wider(names_from = "metric", values_from = c("value", "SES")) %>%
  rename_at(vars(starts_with("value")), ~str_replace(., "value_", "")) %>%
  pivot_longer(-c(rarefyID, rarefied, YEAR, cell, rarefy_resamp), names_to = "metric") %>%
  filter(!is.na(value))

>>>>

Merge rarefied species and fd metrics and null models
```{r}
rarefied_metrics <- full_join(species_metrics, fd_metrics %>% select(-CWM),
                              c("YEAR", "cell", "rarefy_resamp", "rarefyID")) %>%
  full_join(cwm_metrics %>% rename_with(~paste0("CWM_", .), -c("YEAR", "cell", "rarefy_resamp", "rarefyID")), 
            c("YEAR", "cell", "rarefy_resamp", "rarefyID")) %>%
  filter(S != 1) %>% # remove samples that had one species as they result in strange metric values
  select(-qual.FRic)

@@

##	calculate the medians for all the metrics for studies that were rarefied
## exclude categorical variables - they don't make sense as medians
metric_cols <- ungroup(rarefied_metrics) %>%
  select(-c(SamplePool, SampleN, num_years, duration, startYear, endYear, type,
            CWM_diet_5cat, CWM_pelagicspecialist, CWM_diurnal, CWM_nocturnal, 
            CWM_crepuscular, CWM_dir, CWM_lar)) %>%
  tidyr::pivot_longer(-c(rarefyID, rarefied, YEAR, cell, rarefy_resamp), names_to = "metric") %>%
  left_join(null_table %>% select(-c(richness, se, lowerCI, upperCI)),
            by = c("YEAR" = "year", "rarefyID" = "rarefyid", "rarefy_resamp" = "rarefy_resamp", "metric" = "metric"))

@@

# calculate SES and incorporate into metric columns
metric_cols <- metric_cols %>%
  mutate(SES = (value - mean)/sd) %>%
  #for years where species richness = richness of the species pool, FRic sd = 0, so use the original value with no SES correction
  mutate(SES = case_when(
    sd == 0 ~ 0,
    TRUE ~ SES)) %>%
  select(-c(mean, sd)) %>%
  #the next three lines go back to wide format for both SES and unadjusted values, then to long format incorporating SES values
  tidyr::pivot_wider(names_from = "metric", values_from = c("value", "SES")) %>%
  rename_at(vars(starts_with("value")), ~str_replace(., "value_", "")) %>%
  tidyr::pivot_longer(-c(rarefyID, rarefied, YEAR, cell, rarefy_resamp), names_to = "metric") %>%
  filter(!is.na(value))

### 2021-12-16 analysis/05_model_metrics.Rmd

model_data <- metrics %>% 
  filter(metric %in% c("SES_FRic", "SES_FEve", "SES_FDiv", "Jaccard_base", "Jaccard_next",
                       "FRic", "FEve", "FDiv", "S",
                       "bodymass_value"),
         !rarefyID %in% missing_null) %>%
  group_by(metric) %>%
  mutate(study_id =  str_extract(rarefyID, "[^_]+")) %>%
  mutate(year_scaled = scale(year), 
         scale_center = attributes(year_scaled)$`scaled:center`, 
         scale = attributes(year_scaled)$`scaled:scale`) %>%
  ungroup() %>%
  #this study has many low species richness observations, so FD metrics weren't calculated, just drop it
  filter(study_id != 348)

>>>>

model_data <- metrics %>% 
  filter(metric %in% c("SES_FRic", "SES_FEve", "SES_FDiv", "Jaccard_base", "Jaccard_next",
                       "FRic", "FEve", "FDiv", "S",
                       "CWM_bodymass_value"),
         !rarefyID %in% missing_null) %>%
  group_by(metric) %>%
  mutate(study_id =  str_extract(rarefyID, "[^_]+")) %>%
  mutate(year_scaled = scale(year), 
         scale_center = attributes(year_scaled)$`scaled:center`, 
         scale = attributes(year_scaled)$`scaled:scale`) %>%
  ungroup() %>%
  #this study has many low species richness observations, so FD metrics weren't calculated, just drop it
  filter(study_id != 348)
